### 2016-07-25 finalText/results.tex
Final Version of the Text and Short Paper

\chapter{Results}\label{chap:results}
The optimized design of Section~\ref{sec:struct-shared} was synthesized for different network parameters, using Xilinx's Vivado 2015.4 Design Suite, and the synthesis results are reported in Section~\ref{sec:res-synth}, along with their discussion. The performance of the synthesized networks is very promising, showing an approximate $195\times$ speed-up over the LSTM Python model, which was running on a high-end desktop computer with an over clocked CPU. Although the FPGA aims for an Embedded System environment usage, the fairness of this comparison still holds, since a Python/Numpy implementation is slower than a lower-level language implementation (such as in C) and this level of performance would be similar to a C implementation in a low-power, embedded CPU platform. These results are reported in Section~\ref{sec:res-valid}.

\section{Synthesis}\label{sec:res-synth}
In order to assess the impact of the network size on the performance and resource usage of the proposed design, I have synthesized the Verilog description of the proposed network for sizes $N$ of 4, 8, 16 and 32, and for each of these the resource sharing parameter $K_G$ was also varied (as will be explained further ahead, the number of inputs is set to $M=2$). Since performance is the key factor for using an FPGA implementation, the synthesis strategy used was the \verb+Flow_PerfOptimized_high+, and the implementation strategy was \verb+Performance_ExplorePostRoutePhysOpt+ to perform a post-route optimization stage that proved useful in many cases to meet the clock specifications when the first Place\&Route pass failed to meet them. For a network size of 32 and $K_G=8$, the LUT usage exceeded the LUT resources available in the FPGA, so only lower values of $K_G$ were successfully synthesized.

\subsection{Maximum Frequency}\label{sec:res-synth-maxfreq}
Using the synthesis strategy outlined in the previous paragraph, the maximum clock frequency that does not cause timing violations of any sort is plotted in Figure~\ref{fig:maxfreq}. For $N=4$, since there are only 4 rows to be multiplied, the maximum value of $K_G$ is 4, and hence no synthesis was performed for $K_G = 8$; also, since $K_G = 2$ and $N=32$ would use $32(8/2+3) = 224$ DSP slices, that exceeds the 220 slices available in the Zynq 7020, so there is no synthesis data for that value.

First, for all network sizes, it is clear that with increasing $K_G$, the maximum clock speed decreases. This means that there is a critical path in the Matrix-Vector multiplication unit, whose multiplexer becomes increasingly complex for higher values of $K_G$. On the other hand, when $K_G$ is the same, smaller networks are faster than larger networks. The fastest design is an $N=4$ and $K_G = 2$ network, with a clock frequency of $140.154$ MHz, and the slowest one is an $N=32$ and $K_G=4$ network, clocked at $90.498$ MHz. The reference design used for validation in Section~\ref{sec:res-valid} is an $N=8$ and $K_G=2$ network is clocked at $132.89$ MHz, which yields a clock period of $7.525$ ns.

@@

\subsection{DSP Slices Usage}\label{sec:res-synth-dsp}
The estimates made in Equation~\ref{eq:numdsp_network-opt} were proven to be accurate, as Figure~\ref{fig:dspused} confirms. The reference network design uses 56 DSP slices, which corresponds to $25.45\%$ of the total number of DSP slices available.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/dspuse.eps}
    \caption[The number of DSP slices used for several Network Sizes $N$ and resource sharing level $K_G$]{The number of DSP slices used for several Network Sizes $N$ and resource sharing level $K_G$}
    \label{fig:dspused}
\end{figure}

\subsection{Power Usage}\label{sec:res-synth-power}
Another important metric of the performance of a design is its \textbf{power consumption}. All designs yielded a constant baseline value for the \emph{static} power consumption of around 120 mW, and the power consumption reported in Figure~\ref{fig:power} refers to the \emph{total} power consumptions, i.e. both the baseline static power and the dynamic power consumption. It is clear that the smaller the network is, the less power is consumed, as one would expect. Furthermore, an increasing level of resource sharing yields a substantially lower power consumption figure: this is because we use less DSP slices when we increase $K_G$. Of course, even though the power consumption is lower in that case, that comes at the expense of a lower clock frequency and more clock cycles elapsed per forward propagation.

@@

\subsection{Other Resources Usage}\label{sec:res-synth-otheres}
Besides DSP slices, it is also worth mentioning the usage of LUT, LUTRAMs and Flip-Flops. In Table~\ref{tab:lut}, the usage of LUTs is reported, where we can see that although there is not a clear trend on how the LUT usage varies with increasing $K_G$, it is clear, and expectable, that the LUT usage increases with the size of the network by an approximate factor of 2, from $N=2$ to $N=16$. As for $N=32$, the usage does not follow this apparent trend, and rises sharply to 91\%. For $K_G = 8$ the usage even surpasses the maximum amount of LUTs available.
\begin{table}
    \centering
  \begin{tabular}{ | l | c | c | c |}
    \hline
    & $K_G=2$  & $K_G=4$ & $K_G=8$ \\ 
    \hline
    $N=4$ & 6.87\% & 6.04\% & N.A. \\    
    \hline
    $N=8$ & 14.64\% & 13.03\% & 14.11\% \\    
    \hline
    $N=16$ & 28.97\% & 27.72\% & 29.85\% \\    
    \hline
    $N=32$ & N.A. & 91.09\% & N.D. \\    
\hline
  \end{tabular}
    \caption{LUT usage for different $N$ and $K_G$}
    \label{tab:lut}
\end{table}

As for LUT, the FF usage also scales accordingly to a $2\times$ factor. Furthermore the LUTRAM usage scales well and does not pose a limitation on the network size. The usage results are reported in Table~\ref{tab:ff}.

\begin{table}
    \centering
  \begin{tabular}{ | l | c | c |}
    \hline
         & LUTRAM  & FF  \\ 
    \hline
    $N=4$ & 0.18\%  & 3.39\% \\    
    \hline
    $N=8$ & 4.41\%  & 6.7\% \\    
    \hline
    $N=16$ & 8.83\%  & 13.36\% \\    
    \hline
    $N=32$ & 17.66\% &  26.45\% \\    
\hline
  \end{tabular}
    \caption{Flip-flop usage for different $N$ and $K_G$}
    \label{tab:ff}
\end{table}

\section{Validation and Comparison}\label{sec:res-valid}
Over the course of this Section, the functionality of the network is verified against the Python reference module of~\ref{sec:res-pyth} that I have developed as a reference, both for the forward propagation of the network, as well as for the training algorithm. The methodology of this procedure is outlined in Section~\ref{sec:res-meth}. On Section~\ref{sec:res-valid-fpps}, I will assess how the performance results of Section~\ref{sec:res-synth-maxfreq} translate into how many classifications per second this design achieves, and how that figure compares with the capabilities of the Python module.

\subsection{Reference Python Module}\label{sec:res-pyth}
Before performing the Verilog description of the network, I have tested the ideas outlined in Section~\ref{chap:theorBack} first by building a software version of an LSTM Network. Since I wanted to quickly test the ideas without much effort, I decided to use Python and Numpy rather than MATLAB, since the former have higher performance at the same level of code complexity. The code for this Python class is listed in Appendix~\ref{ap1:lstm-pyth}. It supports both the forward propagation of an already trained network, but also provides functions to train a model using the SPSA method outlined in Section~\ref{sec:training_lstm} and also Backpropagation Through-Time as presented in~\cite{Greff15}. As the reader can attest by running the code himself, the SPSA proves to be a valid approach for training these networks, achieving convergence for the Machine Learning task that will be presented in the next paragraph, although it converges slower than Backpropagation Through Time.

@@

\subsection{Methodology}\label{sec:res-meth}
The Python testbench that both trains and verifies the predictions of the software network of Appendix~\ref{ap1:lstm-pyth} is listed in Appendix~\ref{ap1:testbench-pyth}. This script, after achieving convergence, stores the trained network weights in \href{https://docs.python.org/3/library/pickle.html}{Pickle} format. These weights are, then, converted to $Q6.11$ fixed-point format and directly loaded into the Verilog model by the Verilog testbench. The Verilog testbench then applies in order the bits of randomly generated numbers, captures the output bit and checks whether it is correct or not, and counts the number of incorrect bits.

This network has, then, two inputs ($M=2$) and eight neurons ($N=8$). Furthermore, the output of the eight LSTM neurons is reduced to a single output by means of a simple neuron, as the one described in Figure~\ref{fig:modelNeuron_a}. This introduces a further 5 clock cycle overhead that is not accounted in Section~\ref{sec:res-valid-fpps}, since it is not actually part of the LSTM Network and, although it was placed in series in the datapath, after the LSTM network, it can run in parallel with the LSTM network, introducing no overhead whatsoever.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/wave.png}
    \caption[Screenshot of the simulation run of a complete forward propagation for $N=8$ and $K_G=2$]{Screenshot of the simulation run of a complete forward propagation for $N=8$ and $K_G=2$. The period between the two time cursors is the same as the one reported in Table~\ref{tab:process-time}}
    \label{fig:res-meth-wave}
\end{figure}
\subsection{Performance}\label{sec:res-valid-fpps}
To evaluate the performance of the network in terms of how fast it can produce data, I have defined a metric based on how many predictions it can produce per second (i.e. produce a new result bit in the output sequence), in \textbf{millions}. The prediction time is the time elapsed since a new input vector is applied \emph{to} the moment the LSTM Network produces an output vector (as discussed previously, the five clock cycles of the simple neuron at the output are \textbf{not} considered). To do this, we multiply the number of clock cycles yielded by Equation~\ref{eq:numcc_network-opt} by the equivalent clock period from the synthesis clock report of Figure~\ref{fig:maxfreq}. This result is epitomized in Table~\ref{tab:process-time}, where the calculation time of the Python module forward propagation function is also included. This time was measured using the \href{https://docs.python.org/3.5/library/timeit.html}{\textbf{timeit}} module, that allows the evaluation of the execution time of small pieces of code as well as complete functions with arguments. The Python code was run on a Linux System, powered by an i7-3770k Intel Processor, running at 4.2GHz.

\begin{table}
    \centering
    \begin{tabular}{ | l | c | c | c | c | c | }
    \hline
    & $K_G=8$  & $K_G=4$ & $K_G=2$ & Python & Speed-up \\ 
    \hline
    $N=4$ & N.A.  & 360.15 ns  & 292.535 ns & 65 $\mu$s & $\times180$ \\   
    \hline 
    $N=8$ & 960.3 ns  & 494 ns  &  368.725 ns & 72 $\mu$s & $\times194$ \\    
    \hline
    $N=16$ & 1.715 $\mu$s  & 921.5 ns  & 518.05 ns & 96 $\mu$s & $\times185$ \\
    \hline
    $N=32$ & N.D. & 1.78 $\mu$s  & N.A. & 185 $\mu$s  &  $\times104$ \\    
\hline
  \end{tabular}
    \caption{Total processing time for a single forward propagation, for hardware networks of various $K_G$ and the Python software network of equivalent size. The speed-up figure is calculated between the Python processing time versus the best hardware network time of the same size}
    \label{tab:process-time}
\end{table}
The performance increase is impressive, even for the slowest of the designs (the $N=32$ network). The hardware network is, at best, $\times200$ faster than the software counterpart, and at worst $\times100$ faster. Also, it is noticeable that increasing the level of resource sharing increases the computation time, since the level of parallelism is lower.

 Since the previous values are the time needed for a \emph{single} forward propagation, to know how many forward propagations we can perform \emph{per second}, we only need to invert the previous values. These values are presented in Figure~\ref{fig:Mclass-psec}. While the $N=8$ and $K_G=2$ network is able to perform around 3.2 million predictions per second, the Python model can only output around 14 thousand predictions, which is a very significant result that proves the relevance of this implementation.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/Mclass-psec.eps}
    \caption[Millions of classifications per second of each design according to the network size $N$]{Millions of classifications per second of each design according to the network size $N$. The comparison is between the software Python model and 3 networks of different levels of resource sharing $K_G$.}
    \label{fig:Mclass-psec}
\end{figure}

>>>>

\chapter{Results}\label{chap:results}
The optimized design of Section~\ref{sec:struct-shared} was synthesized for different network parameters, using Xilinx's Vivado 2015.4 Design Suite,
and the synthesis results are reported in Section~\ref{sec:res-synth}, along with their discussion. The performance of the synthesized networks is very promising,
showing an approximate $251\times$ speed-up over the LSTM Python model, which was running on a high-end desktop computer with an overclocked CPU, and a $14\times$ speed-up
over the most recent hardware implementation~\cite{Chang15} Although the FPGA aims for an Embedded System environment usage, the fairness of the comparison with a Desktop computer
still holds, since a Python/Numpy implementation is slower than a lower-level language implementation (such as in C) and this level of performance would be similar to a C implementation
in a low-power, embedded CPU platform. These results are reported in Section~\ref{sec:res-valid}.

\section{Synthesis}\label{sec:res-synth}
The proposed network was first synthesized aiming a Xilinx XC7Z020 SoC, for sizes $N$ of 4, 8, 16 and 32, varying the resource sharing parameter $K_G$, and the number of inputs set
to $M=2$. For a network size of 32 and $K_G=8$, the LUT usage exceeded the LUT resources available in the FPGA, so only lower values of $K_G$ were
successfully synthesized. This is because of the complexity of the input multiplexer of the Matrix-vector Dot Product Module in~\ref{sec:dotprod_sec}. To synthesize the
design for sizes $N$ of 64, 128 and 256, a Virtex-7 VC707 board was used.

In order to assess the impact of the network size on the performance and resource usage of the proposed design, the Verilog description of the proposed network was first synthesized
aiming a Xilinx XC7Z020 SoC platform, for sizes $N \in \left\{4, 8, 16, 32 \right\}$, and for each of these the resource sharing parameter $K_G$ was also varied (as will be explained further ahead,
the number of inputs is set to $M=2$). Since performance is the key factor for using an FPGA implementation, the synthesis strategy used was the \verb+Flow_PerfOptimized_high+, and the implementation
strategy was \verb+Performance_ExplorePostRoutePhysOpt+ to perform a post-route optimization stage that proved useful in many cases to meet the clock specifications when the first Place\&Route pass
failed to meet them. For a network size of 32 and $K_G=8$, the LUT usage exceeded the LUT resources available in the FPGA, so only lower values of $K_G$ were successfully synthesized (N.D. in the tables).
Networks of sizes $N \in \left\{64, 128\right\}$ were also synthesized for a Virtex-7 VC707 board.

\subsection{Maximum Frequency}\label{sec:res-synth-maxfreq}
Using the synthesis strategy outlined in the previous paragraph, the maximum clock frequency that does not cause timing violations of any sort is plotted in
Figure~\ref{fig:maxfreq}. For $N=4$, since there are only 4 rows to be multiplied, the maximum value of $K_G$ is 4, and hence no synthesis was performed for $K_G = 8$ (N.A. in the tables);
also, since $K_G = 2$ and $N=32$ would use $32(8/2+3) = 224$ DSP slices, that exceeds the 220 slices available in the Zynq 7020, so there is no synthesis data for that value.

First, for all network sizes, it is clear that with increasing $K_G$, the maximum clock speed decreases. This means that there is a critical path in the Matrix-Vector multiplication unit,
whose multiplexer becomes increasingly complex for higher values of $K_G$. On the other hand, when $K_G$ is the same, smaller networks are faster than larger networks. The fastest design
is an $N=4$ and $K_G = 2$ network, with a clock frequency of $158.228$ MHz, and the slowest one is an $N=32$ and $K_G=4$ network, clocked at $101.523$ MHz. The reference design used for
validation in Section~\ref{sec:res-valid} is an $N=8$ and $K_G=2$ network is clocked at $154.321$ MHz, which yields a clock period of $6.48$ ns.

@@

\subsection{DSP Slices Usage}\label{sec:res-synth-dsp}
The estimates made in Equation~\ref{eq:numdsp_network-opt} were proven to be accurate, as Figure~\ref{fig:dspused} confirms. The reference network design uses 56 DSP slices, which
corresponds to $25.45\%$ of the total number of DSP slices available.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/dspuse.eps}
    \caption[The number of DSP slices used for several Network Sizes $N$ and resource sharing level $K_G$]{The number of DSP slices used for several Network Sizes $N$ and resource sharing level $K_G$}
    \label{fig:dspused}
\end{figure}

\subsection{Power Consumption}\label{sec:res-synth-power}
Another important metric of the performance of a design is its \textbf{power consumption}. All designs yielded a constant baseline value for the \emph{static} power consumption of around 120 mW,
and the power consumption reported in Figure~\ref{fig:power} refers to the \emph{total} power consumptions, i.e. both the baseline static power and the dynamic power consumption.
It is clear that the smaller the network is, the less power is consumed, as one would expect. Furthermore, an increasing level of resource sharing, $K_G$, yields a substantially lower
power consumption figure: this is because less DSP slices are used when $K_G$ is increased. Of course, even though the power consumption is lower in that case, that comes at the
expense of a lower clock frequency and a higher computation overhead per forward propagation.

@@

\subsection{Other Resources Usage}\label{sec:res-synth-otheres}
The LUT, LUTRAM and Flip-Flop usages are discussed here. In Table~\ref{tab:lut}, the usage of LUTs is reported, where we can see
that although there is not a clear trend on how the LUT usage varies with increasing $K_G$, it is clear, and expectable, that
the LUT usage increases with the size of the network by an approximate factor of 2, from $N=2$ to $N=16$. As for $N=32$, the
usage does not follow this apparent trend, and rises sharply to 91\%. For $K_G = 8$ the usage surpasses the maximum amount
of LUTs available in the XC7Z020.

\begin{table}

    \centering
  \begin{tabular}{ | l | c | c | c |}
    \hline
    & $K_G=2$  & $K_G=4$ & $K_G=8$ \\
    \hline
    $N=4$ & 6.87\% & 6.04\% & N.A. \\
    \hline
    $N=8$ & 14.64\% & 13.03\% & 14.11\% \\
    \hline
    $N=16$ & 28.97\% & 27.72\% & 29.85\% \\
    \hline
    $N=32$ & N.A. & 91.09\% & N.D. \\
\hline
  \end{tabular}
  \caption{LUT usage for different $N$ and $K_G$ in the XC7Z020}
  \label{tab:lut}
\end{table}

As for LUTs, the FF usage also scales accordingly to a $2\times$ factor. In terms of LUTRAM, used to store the network weights,
the amount used increases by 2 with increasing values of $N$, as before, and \textbf{does not} depend on $K_G$. This is because
the amount of weights only depends on the network sizes $M$ and $N$, and not on $K_G$. Furthermore, unlike LUTs, it scales well
with increasing network sizes, and does not pose a limitation on the network size. The usage results for the XC7Z020 are reported in Table~\ref{tab:ff},
and in Table~\ref{tab:ff-virtx7} the usages for the VC707.

\begin{table}
    \centering
  \begin{tabular}{ | l | c | c |}
    \hline
         & LUTRAM  & FF  \\
    \hline
    $N=4$ & 0.18\%  & 3.39\% \\
    \hline
    $N=8$ & 4.41\%  & 6.7\% \\
    \hline
    $N=16$ & 8.83\%  & 13.36\% \\
    \hline
    $N=32$ & 17.66\% &  26.45\% \\
\hline
  \end{tabular}
  \caption{Flip-flop and LUTRAM usage for the XC7Z020}
	\label{tab:ff}
\end{table}

\begin{table}

    \centering
  \begin{tabular}{ | l | c | c | c |}
    \hline
         & LUTRAM  & FF & LUT \\
    \hline
    $N=64$ & 24.22\%  & 9.14\% & 24.22\% \\
    \hline
    $N=128$ & 14.09\%  & 18.3\% & 41.82\%\\
    \hline
  \end{tabular}
  \caption{Flip-flop, LUTRAM and LUT usage for the VC707}
  \label{tab:ff-virtx7}
\end{table}

\section{Validation and Comparison}\label{sec:res-valid}
Over the course of this Section, the functionality of the network is verified against the Python reference module of~\ref{sec:res-pyth} that was developed as a
reference, both for the forward propagation of the network, as well as for the training algorithm. The methodology of this procedure is outlined in Section~\ref{sec:res-meth}.
On Section~\ref{sec:res-valid-fpps}, the performance results of Section~\ref{sec:res-synth-maxfreq} is translated into how many classifications per second this design
achieves, and how that figure compares with the capabilities of the Python module and the current existing hardware implementation~\cite{Chang15}.

\subsection{Reference Python Module}\label{sec:res-pyth}
Before performing the Verilog description of the network, the ideas outlined in Section~\ref{chap:theorBack} were tested first by building a software version of an LSTM Network.
Since we want to quickly test the ideas without much effort, Python and Numpy were used rather than MATLAB, since Python/Numpy has higher performance at the same level of
code complexity. The code for this Python class is listed in Appendix~\ref{ap1:lstm-pyth}. It supports both the forward propagation of an already trained network, but also provides
functions to train a model using the SPSA method outlined in Section~\ref{sec:training_lstm} and also Backpropagation Through-Time as presented in~\cite{Greff15}. As the reader can attest
by running the code himself, the SPSA proves to be a valid approach for training these networks, achieving convergence for the Machine Learning task that will be presented in the next paragraph,
although it converges slower than Backpropagation Through Time.

@@

\subsection{Methodology}\label{sec:res-meth}
The Python testbench that both trains and verifies the predictions of the software network of Appendix~\ref{ap1:lstm-pyth} is listed in Appendix~\ref{ap1:testbench-pyth}. The scripts train the
network for a maximum of 100 epochs, and in each epoch, the network is presented with 5000 8 bit additions for training, and then 100 8 bit additions for testing, where the prediction error is
calculated, and the training convergence is verified (when there is zero error for two consecutive times). After achieving convergence, the script stores the trained network weights in
\href{https://docs.python.org/3/library/pickle.html}{Pickle} format. These weights are, then, converted to $Q6.11$ fixed-point format and directly loaded into the Verilog model by the Verilog
testbench. The Verilog testbench then applies in order the bits of randomly generated numbers, captures the output bit and checks
whether it is correct or not, and counts the number of incorrect bits.

This network has, then, two inputs ($M=2$) and eight neurons ($N=8$). Furthermore, the output of the eight LSTM neurons is reduced to a single output by means of a simple neuron, as the one
described in Figure~\ref{fig:modelNeuron_a}. This introduces a further 5 clock cycle overhead that is not accounted in Section~\ref{sec:res-valid-fpps}, since it is not actually part of the
LSTM Network. Although it was placed in series in the datapath, after the LSTM network, it can be run in parallel with the LSTM network, introducing no overhead whatsoever.

@@

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\linewidth]{figures/wave.png}
    \caption[Screenshot of the simulation run of a complete forward propagation for $N=8$ and $K_G=2$]{Screenshot of the simulation run of a complete forward propagation for $N=8$ and $K_G=2$. The period between the two time cursors is the same as the one reported in Table~\ref{tab:process-time}}
    \label{fig:res-meth-wave}
\end{sidewaysfigure}

\subsection{Performance}\label{sec:res-valid-fpps}
To evaluate the performance, a metric was defined based on how many predictions it can produce per second (i.e. produce a new result bit in the output sequence),
in \textbf{millions}. The metric on~\cite{Chang15} is not very informative, since
although a system can perform many calculations per second, if those operations are redundant, that metric has no relevant information
regarding \textbf{how fast} the system can \textbf{perform the actual task it was meant to do}, which in this case is a complete Forward
Propagation through the network. The prediction time is the time elapsed since a new input vector is applied \emph{to} the moment the LSTM
Network produces an output vector. Hence, we multiply the number of clock cycles yielded by Equation~\ref{eq:numcc_network-opt} by the equivalent clock period
from the synthesis clock report of Figure~\ref{fig:maxfreq}.
This result is epitomized in Table~\ref{tab:process-time}, where the calculation time of the Python module's Forward Propagation function is
also included. This time was measured using the \textbf{timeit}\footnote{https://docs.python.org/3.5/library/timeit.html}  module, that allows
the evaluation of the execution time of small pieces of code as well as complete functions with arguments. The Python code was run on a Linux System,
powered by an i7-3770k Intel Processor, running at 4.2GHz.

\begin{table}
    \centering
    \begin{tabular}{ | l | c | c | c | c | c | }
    \hline
    & $K_G=8$  & $K_G=4$ & $K_G=2$ & Python & Speed-up \\
    \hline
    $N=4$ & N.A.  & 309.68 ns  & 259.12 ns & 65 $\mu$s & $\times251$ \\
    \hline
    $N=8$ & 793.46 ns  & 421.12 ns  &  317.52 ns & 72 $\mu$s & $\times228$ \\
    \hline
    $N=16$ & 1.497 $\mu$s  & 738.19 ns  & 461.336 ns & 96 $\mu$s & $\times208$ \\
    \hline
    $N=32$ & N.D.          & 1.586 $\mu$s & N.A.       & 185 $\mu$s  &  $\times117$ \\
		\hline
  \end{tabular}
  \caption{Total processing time for a single forward propagation on the XC7Z020}
  \label{tab:process-time}
\end{table}
The performance increase is significant, even for the slowest of the designs (the $N=32$ network). The hardware network is, at best, $\times251$ faster than the software counterpart,
and at worst $\times117$ faster. Also, it is noticeable that increasing the level of resource sharing increases the computation time, since the level of parallelism is lower.

Since the previous values are the time needed for a \emph{single} forward propagation, to know how many forward propagations we can perform \emph{per second},
we only need to invert the previous values. These values are presented in Figure~\ref{fig:Mclass-psec}. While the $N=8$ and $K_G=2$ network is able to
perform around 3.15 million predictions per second, the Python model can only output around 14 thousand predictions, which is a very significant result that
proves the relevance of this implementation.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/Mclass-psec.eps}
    \caption[Millions of classifications per second of each design according to the network size $N$]{Millions of classifications per second of each design according to the network size $N$. The comparison is between the software Python model and 3 networks of different levels of resource sharing $K_G$.}
    \label{fig:Mclass-psec}
\end{figure}

As for the larger-sized networks synthesized in the VC707, the results are also very promising. For network sizes of $N=64$ and $N=128$, a complete
forward propagation takes 1.14 $\mu$s and 2.052 $\mu$s respectively, and for both the maximum clock frequency achievable was 140.845 MHz. Since the
design~\cite{Chang15}, for $N=128$, takes an estimated 29.13 $\mu$s, this design yields an improvement of $14\times$ over it.

### 2016-07-25 finalText/stateArt.tex
Final Version of the Text and Short Paper

\begin{itemize}
    \item \textbf{Handwriting Recognition} -- an LSTM-based network~\cite{Bertolami09}, submitted by a team of the Technische Universität München, won the 2009 ICDAR Handwriting Recognition Contest, achieveing a recognition rate of up to 91\%~\cite{ICDAR09}. LSTMs have also proven to surpass HMM-based models in terms of optical character recognition of printed text, as~\cite{Breuel13} suggests.

@@

\section{Hardware Implementations of LSTM}\label{sec:sa_hardware}
Hitherto, there is but one actual implementation of an LSTM network in hardware, published recently (November 2015) by Chang et al.~\cite{Chang15} in the Computing Research Repository (CoRR). It consists on the proposal of an LSTM cell architecture for dedicated hardware, targeting a Xilinx\textregistered~ Zedboard implementation. It uses a character-level language model from Andrej Karpathy~\footnote{\href{https://github.com/karpathy/char-rnn}{https://github.com/karpathy/char-rnn}}, written in Lua using the Torch7 framework~\footnote{\href{http://torch.ch/}{http://torch.ch/}} (the Lua calls are implemented in C, so no performance is lost).

Although the article is not clear on whether there is active learning by the ARM CPU -- the authors refer that the CPU loads the weights before operation, and that it changes them during operation, although how and why that change is done is not even clearly explained, neither mathematically nor conceptually --, \textbf{there is no on-chip learning module} in the FPGA according to the description provided.

The implementation itself makes an extensive use of the Multiply-and-Accumulate units (MAC) of the FPGA, which since they are limited in number in the target platform, limits the number of neurons that we can deploy in parallel. The authors report a nearly 75\% usage for only 8 LSTM neurons. Although an apparently excessive number, I am left to scrutinize whether the usage of MACs is compulsory to obtain a better performance than a CPU or if it can be discarded, allowing for smaller cells. This way, we can have a layer of more neurons occupying the same area and, hopefully, the same, or even fewer, resources within the FPGA.

\section{Training Algorithms and Implementation}\label{sec:sa_training}
As stated in~\ref{sec:sa_hardware}, the work of~\cite{Chang15} does not feature on-chip learning at FPGA level, although there are a handful proposed solutions for it in recent literature. I will particularly look into the ones that use SPSA (see Section~\ref{sec:training_lstm}), since that is the training algorithm of choice for my proposed solution, and also to the ones that particularly apply SPSA to the training of Neural Networks.

The SPSA algorithm was initially published by Spall in~\cite{Spall98}, and its theoretical details are outlined in Section~\ref{sec:training_lstm}. As to its applications to the training of general Neural Networks, the earliest examples come from 1995 and 1996 in~\cite{Maeda95, Cauwen96} where SPSA is used to train a VLSI \emph{Analog} Neural Networks, a time where the memory resources of digital circuitry were limited, and so most of these structures were analog-based. Its adequacy was also established for \textbf{control problems}, such as those proposed in~\cite{Figueiredo97}, where a Time Delay Neural Network is used to control an unknown plant in a linear feedback system.

In 2005, Maeda and Wakamura published a proposed SPSA hardware implementation~\cite{Maeda05} to train an Hopfield Neural Network in an FPGA (and thus a digital system), achieving promising results in an Altera FPGA. The article carefully delineates the approach taken, and also the hardware architecture designed, so it is a very good reference for the design that I will have to implement.

@@

\section{Final Overview}\label{sec:overview}
As we could attest from this small literature survey, although there is already an hardware implementation of LSTM, there is still a good deal of room for improvement by adding on chip-learning to the system, and also to restrict it to a smaller use of the FPGA resources, allowing it to accomodate a more complex network. Furthermore,~\cite{Tavear13} shows that, at least for that particular case, the LSTM network doesn't suffer a great performance impact from using SPSA training, as opposed to the more common BPTT, and~\cite{Maeda05} showed that an SPSA hardware implementation is feasible. These three conclusions combined indicate that the idea of a hardware LSTM network with on-chip learning using SPSA is viable.

Besides the proposed hardware implementation, it would be advantageous to perform \emph{benchmarking} to see how well it compares with software solutions, and Section~\ref{sec:sa_apps} suggests a handful of examples to test my final solution.

>>>>

\begin{itemize}
    \item \textbf{Handwriting Recognition} -- an LSTM-based network~\cite{Bertolami09}, submitted by a team of the Technische Universität München, won the 2009 ICDAR Handwriting Recognition Contest, achieving a recognition rate of up to 91\%~\cite{ICDAR09}. LSTMs have also proven to surpass HMM-based models in terms of optical character recognition of printed text, as~\cite{Breuel13} suggests.

@@

\section{Hardware Implementations of LSTM}\label{sec:sa_hardware}
Hitherto, there is but one actual implementation of an LSTM network in hardware, published recently
(Januray 2016) by Chang et al.~\cite{Chang15} in the international Journal of Advanced Research in
Electrical, Electronics and Instrumentation Engineering. It consists on the proposal of an LSTM network with $N=128$ for dedicated hardware,
targeting a Xilinx\textregistered~ Zedboard implementation. It uses a character-level language model from Andrej
Karpathy~\footnote{\href{https://github.com/karpathy/char-rnn}{https://github.com/karpathy/char-rnn}}, written in
Lua using the Torch7 framework~\footnote{\href{http://torch.ch/}{http://torch.ch/}}.

Although the article is not clear on whether there is active learning by the ARM CPU -- the authors refer that the CPU
loads the weights before operation, and that it changes them during operation, although how and why that change is done
is not even clearly explained, neither mathematically nor conceptually --, \textbf{there is no on-chip learning module}
in the FPGA according to the description provided. For that reason, the system stores the weights in external memory,
and accesses them using Direct Memory Access (DMA). Furthermore, the level of parallelism is low, since for an
$N=128$ network, the design uses only $22.73\%$ of the DSP48's available in the board. This is because the matrix multiplications
are serialized, and only one DSP is used for each matrix-vector product (aside from other uses).

The implementation is not flexible, since there is nothing that suggests a parameterization of the HDL description.
Hence, it is only possible to synthesize networks of size $N=128$, meaning that there needs to be a redesign of the network
for other network sizes. On the other hand, this is a valid approach for special applications that require networks of
this size, and where we only need a co-processor for LSTM networks, and not a dedicated standalone system.

\section{Training Algorithms and Implementation}\label{sec:sa_training}
As stated in~\ref{sec:sa_hardware}, the work of~\cite{Chang15} does not feature on-chip learning at FPGA level, although there
are a handful proposed solutions for it in recent literature. I will particularly look into the ones that use SPSA (see
Section~\ref{sec:training_lstm}), since that is the training algorithm of choice for my proposed solution, and also to the
ones that particularly apply SPSA to the training of Neural Networks~\cite{Maeda05}.

The SPSA algorithm was initially published by Spall in~\cite{Spall98}, and its theoretical details, are outlined in
Section~\ref{sec:training_lstm}. As to its applications to the training of general Neural Networks, the earliest examples
come from 1995 and 1996 in~\cite{Maeda95, Cauwen96} where SPSA is used to train a VLSI \emph{Analog} Neural Networks, a time
where the memory resources of digital circuitry were limited, and so most of these structures were analog-based. Its adequacy
was also established for \textbf{control problems}, such as those proposed in~\cite{Figueiredo97}, where a Time Delay Neural
Network is used to control an unknown plant in a linear feedback system.

In 2005, Maeda and Wakamura published a proposed SPSA hardware implementation~\cite{Maeda05} to train an Hopfield Neural Network
in an FPGA (and thus a digital system), achieving promising results in an Altera FPGA. The article carefully delineates the approach
taken, and also the hardware architecture designed.

@@

\section{Final Overview}\label{sec:overview}
As we could attest from this small literature survey, although there is already an hardware implementation of LSTM, there is still a good deal
of room for improvement by adding on chip-learning to the system, and also to improve its performance and flexibility.
Furthermore,~\cite{Tavear13} shows that, at least for that particular case, the LSTM network doesn't suffer a great performance impact from
using SPSA training, as opposed to the more common BPTT, and~\cite{Maeda05} showed that an
SPSA hardware implementation is feasible.
